---
title: "week 4 project"
author: "Scott Rodgers"
date: "July 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, include=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(rattle)
library(parallel)
library(doParallel)
library(dplyr)
cluster <- makeCluster(detectCores() - 2) 
registerDoParallel(cluster)

```

## Read in and clean data
### This code reads in 2 csvs, reduces the data frame to the 53 features(removing columns that only have na's and cols not in the test set). Lastly the function turns classe into a factor and creates a validation data set from the training data.

```{r read data,cache=TRUE}
training <- read.csv("C:/Users/Srodger/Documents/R/coursera/ml/week 4 project/pml-training.csv")
testing  <- read.csv("C:/Users/Srodger/Documents/R/coursera/ml/week 4 project/pml-testing.csv")
set.seed(222)

testing[,7:159] <- sapply(testing[,7:159],as.numeric) 
testing <- testing %>% select(roll_belt:problem_id) #%>% sample_frac(.4)
testing <- testing %>% select_if(~ !any(is.na(.)))

training[,7:159] <- sapply(training[,7:159],as.numeric) 
training <- training %>% select(roll_belt:classe) #%>% sample_frac(.4)
cols <- colnames(testing)
cols[53] <- "classe"
training  <- training %>%select(cols)

training <- training %>% select_if(~ !any(is.na(.)))
training$classe <- as.factor(training$classe)

inTrain <-createDataPartition(y=training$classe, p=0.6, list=FALSE)
training <- training[inTrain,]
validation <- training[-inTrain,]
```

### This section of code sets the train control for the caret train function to use 5 fold cross validation (this will be used for all models), enables parrallel processing and does not return data to increase performance. Lastly this function inputes any missing values for using k nearest neighbor.

```{r control and impute, cache=TRUE}

fitControl <- trainControl(method = "cv",
                           number = 5,
                           allowParallel = TRUE, 
                           returnData = FALSE)



train_pre_obj <- preProcess(training, method = "knnImpute")
imputed_train <- predict(train_pre_obj, training)

validation <- predict(train_pre_obj, validation)

```


### Build decsion tree, display tree, and calulate confusion matrix and statistics.
```{r build d tree, warning=FALSE, cache=TRUE}
fit_dt <- train(classe~., method="rpart",data=training, trControl = fitControl)
fancyRpartPlot(fit_dt$finalModel)
pred_dt <- predict(fit_dt, validation)
confusionMatrix(pred_dt, validation$classe)
```

### Build Random Forest model, show most important variables, and calculate confusion matrix. Lastly plot fit and final model.

```{r build RF}
# Random Forest Run previously loaded using readRDS
#fit_rf <- train(classe ~., method="rf", data=imputed_train, trControl = fitControl)
#saveRDS(fit_rf, "randf.RData")

fit_rf <- readRDS("C:/Users/Srodger/Documents/R/coursera/ml/week 4 project/randf.RData")
caret::varImp(fit_rf)

pred_rf <- predict(fit_rf, validation)
confusionMatrix(pred_rf, validation$classe)
plot(fit_rf, main="Cross Validation shows 27 predictors gives max accuracy")
plot(fit_rf$finalModel, main ="Final Model showing < 100 trees for best ROI (time vs accuracy)")
```


### Build boosted model, caluclate confusion matrix and plot fit
```{r boosting, cache=TRUE}

# Boosted Tree Run previously loaded using readRDS
#fit_bt <- train(classe ~., method="gbm", data=imputed_train, trControl = fitControl)
#saveRDS(fit_bt, "boostedtree.RData")

fit_bt <- readRDS("C:/Users/Srodger/Documents/R/coursera/ml/week 4 project/boostedtree.RData")

pred_bt <- predict(fit_bt, validation)
confusionMatrix(pred_bt, validation$classe)
plot(fit_bt, main=" Highest Accuracy comes from a tree depth of 3 and >140 boosting iterations")

```

### Compare all three models
```{r compare models}
# load the library

results <- resamples(list(DecisionTree=fit_dt, RandomForest=fit_rf, GBM=fit_bt))
summary(results)
dotplot(results)
```

## Random Forest model is selected for highest accuracy and kappa. Predicting test set data using random forest model created:
```{r Selected Model predictions}
# impute values
testing <- predict(train_pre_obj, testing)

predF <-predict(fit_rf, testing)
predF

```

